---
title: "BSMM-lab-2"
subtitle: "BSMM 8740 Fall 2023"
author: "Manushi Dalal"
date: "September 25, 2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r}
#install.packages("tidyverse")
#install.packages("gt")
#install.packages("gtExtras")     
#install.packages("DataExplorer")
#install.packages("tidymodels")
#install.packages("magrittr")
```

```{r}
library(magrittr)
library(tidyverse)
library(tidymodels)
library(gt)
library(gtExtras)
library(DataExplorer)
```

```{r load-pkg-data}
#| message: false
the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```

## Exercises

### Exercise 1

```{r}
glimpse(the_tate_artists)
```

```{r}
View(the_tate)
View(the_tate_artists)
```

```{r}
glimpse(the_tate)
```

```{r}
summary(the_tate)
```

```{r}
the_tate |> 
  dplyr::group_by(the_tate$artistId) |> 
  dplyr::summarize(N=n(),min_yearworking= min(the_tate$year,na.rm = TRUE),
                   max_yearworking= max(the_tate$year,na.rm = TRUE),
                   min_aquistionyear=min(the_tate$acquisitionYear, na.rm = TRUE),
                   max_acquisitionyear=max(the_tate$year,na.rm = TRUE)  )
```

```{r}
library(dplyr)
# Calculate the number of unique artists using select
unique_artists_count <- the_tate |>
  select(artistId) |>
  distinct() |>
  nrow()
```

```{r}
unique_artists_count
```

```{r}
# Find the range of years artists worked
min_year_worked <- min(the_tate$year,na.rm = TRUE)
max_year_worked <- max(the_tate$year,na.rm = TRUE)
```

```{r}
# Find the range of acquisition years
min_acquisition_year <- min(the_tate$acquisitionYear, na.rm = TRUE)
max_acquisition_year <- max(the_tate$acquisitionYear, na.rm = TRUE)
```

```{r}
DataExplorer::introduce(the_tate)
```

```{r}
DataExplorer::introduce(the_tate_artists)
```

```{r}
DataExplorer::plot_missing(the_tate)
```

```{r}
DataExplorer::plot_missing(the_tate_artists)
```

The the_tate dataset has **3342** unique artists who worked from **1545** to **2012**. The works were acquired between the years **1823** and **2013**.

### Exercise 2

```{r}
works_with_missing_years <- sum(is.na(the_tate$year))
works_with_missing_years
```

How number of works with missing dates is **5397**.

```{r}
the_tate|>
  dplyr::filter(is.na(year))|>
  dplyr::distinct(artist)
```

```{r}
artists_with_missing_dates <- the_tate |>
  filter(is.na(year)) |>
  distinct(artist) |>
  nrow()
```

```{r}
artists_with_missing_dates
```

The number of artists whose works have missing dates is **461**.

```{r}
# Count the number of works missing years for each artist
artist_missing_year_counts <- the_tate|>
  dplyr::filter(is.na(year)) |>
  group_by(artist) |>
  summarise(missing_years = sum(is.na(year))) |>
  arrange(desc(missing_years)) |>
  as_tibble()
```

```{r}
# Determine how many artists have works with missing years
artists_with_missing_years <- nrow(artist_missing_year_counts)
artists_with_missing_years
```

```{r}
# Calculate the percent of total missing data for each artist
artist_missing_year_counts <- artist_missing_year_counts |>
  mutate(percentage = (missing_years / works_with_missing_years) * 100)
```

```{r}
# Calculate the cumulative percent of missing data
artist_missing_year_counts <- artist_missing_year_counts |>
  mutate(cumulative_percentage = cumsum(percentage))
```

```{r}
# Identify the smallest number of artists needed to resolve at least 50% of the missing year data
artists_to_resolve_50_percent <- min(which(artist_missing_year_counts$cumulative_percentage >= 50))

artists_to_resolve_50_percent
```

It would require resolving missing year data for only **11** artists resolve resolve at least 50% of the missing data.

The missing year data likely to be classified as **MAR**.

### Exercise 3

```{r}
# Group the data by artist and count the number of works for each artist
artist_work_counts <- the_tate |>
  group_by(artist) |>
  summarize(title = n()) |>
  arrange(desc(title))

# Display the top 10 artists by the number of works
top_10_artists <- artist_work_counts |>
  slice_head(n = 10)
```

```{r}
View(artist_work_counts)
View(top_10_artists)
```

The artist with the most works in the Tate collection is **Turner, Joseph Mallord William**.

The artist with the tenth-most works in the Tate collection is **Warhol, Andy**.

### Exercise 4

```{r}
total_works <- nrow(the_tate)
artist_work_counts %<>%mutate(percentage=(title/total_works)*100)
```

```{r}
View(artist_work_counts)
```

```{r}
# Create the table using gt
table <- artist_work_counts |>
  gt() |>
  fmt_number(
    columns = c(title, percentage), # Format both title and percentage columns
    decimals = 2 # No decimal places for title, and decimals for percentage
  ) |>
  tab_header(title = "Top Artists by Number of Works and Percentage of Collection")

# Print the formatted table
print(table)
```

The artist with the greatest number of works in the Tate collection represent \_\_\_% of the total number of works

### Exercise 5

```{r}
total_rows <- total_works

# Select only the columns for artist and title, then count distinct pairs
distinct_artist_title_pair <- the_tate |> 
  select(artist,title) |> 
  distinct()
```

```{r}
# Count of distinct artist-title pairs
distinct_count <- nrow(distinct_artist_title_pair)
total_rows
```

```{r}
distinct_count
```

```{r}
# Count the number of duplicated artist-title pairs
duplicated_count <- total_rows - distinct_count
cat("Duplicated artist-title pairs:", duplicated_count, "\n")
```

There are **23705** duplicate artist-title pairs

### Exercise 6

```{r}
# Load the required libraries
library(dplyr)
library(tidyr)

# Calculate the area of each artwork and add it as a new column
the_tate <- the_tate %>%
  mutate(area_cm2 = width * height)

# Select artist, title, and area, remove NA values
selected_artworks <- the_tate %>%
  select(artist, title, area_cm2) %>%
  drop_na()  # Remove rows with NA values

# Order the works by area
ordered_artworks <- selected_artworks %>%
  arrange(area_cm2)

# Find the largest artwork in the collection
largest_artwork <- ordered_artworks %>%
  slice_tail(n = 1)

# Find the smallest artwork in the collection
smallest_artwork <- ordered_artworks %>%
  slice_head(n = 1)

# Print the largest and smallest artworks

print(largest_artwork)
```

```{r}
print(smallest_artwork)
```

The artist with the largest work in the tate collection is **Therrien, Robert**

The artist with the smallest work in the collection is **Mesens, E.L.T.**. The smallest work has area 237 $\text{cm}^2$

### Exercise 7

```{r}
view(the_tate_artists)
```

```{r}
# Load the required libraries
library(dplyr)



# Left join the tables and group the result by gender
gender_grouped<- the_tate |>
  left_join(the_tate_artists, by = c("artist" = "name")) |>  filter(!is.na(gender)) |> 
  group_by(gender) 
```

```{r}
# Show the resulting table
gender_grouped
```

### Exercise 8

```{r}
library(readr)
dataofstock <- read_csv("data/SPX_HistoricalData_1692322132002.csv")
```

```{r}
View(dataofstock)
```

```{r}
# Add a column for the year of the transaction
dataofstock <- dataofstock |>
  mutate(Year = lubridate::year(as.Date(Date, format = "%m/%d/%Y")))
```

```{r}
dataofstock <- dataofstock |>
  rename("close"=`Close/Last`)
dataofstock <- dataofstock |>
  mutate(rd = log(lead(close) / close))
dataofstock <- dataofstock |>
  mutate(vard = rd^2)
```

```{r}
summary_data <- dataofstock |>
  group_by(Year) |>
  summarize(
    Annual_Return = (exp(sum(rd, na.rm = TRUE)) - 1)*100,
    Annual_StdDev = sqrt(sum(vard, na.rm = TRUE))*100,
    .groups = "drop"
  )

# Print the summary data
print(summary_data)
```

The annual return in the SPX price in 2020 was **-13.98**%.

The corresponding price volatility was **34.70**%.

### Exercise 9

```{r}
# Calculate the period volatility as the standard deviation of annual returns
period_volatility <- sd(summary_data$Annual_Return)

# Print the period volatility
cat("The period volatility was:", round(period_volatility, 2), "%\n")
```

The period volatility was **19.52**%

### 
